Certified Knowledge Compilation \\ with Application to Verified Model Counting

### 2023-03-15 doc/mc-data.tex
Lots of updates

\newcommand{\progname}[1]{\textsc{#1}}
\newcommand{\dfour}{\progname{d4}}
\newcommand{\Dfour}{\progname{d4}}
\newcommand{\generator}{\progname{d2p}}
\newcommand{\Generator}{\progname{d2p}}
\newcommand{\checker}{\progname{crat-check}}
\newcommand{\Checker}{\progname{Crat-check}}

@@

@@

\begin{figure}
\centering{%
\begin{tikzpicture}[scale = 0.90]
  \begin{axis}[mark options={scale=0.55},grid=both, grid style={black!10}, ymode=log, ymin=0.01, ymax=20000,
      legend style={at={(0.30,0.98)}},
      legend cell align={left},
                              x post scale=1.8, y post scale=2.4,
                              xmode=log,xmin=0.01,xmax=4000,
                              xtick={0.01, 0.1,1.0,10,100,1000,10000}, xticklabels={0.01, 0.1, 1.0, 10, 100, {1,000}, {10,000}},
                              ytick={0.01, 0.1,1.0,10,100,1000,10000,100000}, yticklabels={0.01, 0.1, 1.0, 10, 100, {1,000},{10,000},{100,000}},
                              xlabel={D4 runtime (seconds)}, ylabel={CRAT generation and checking runtime (seconds)},
%                              title={D4 Defining Clause Generation}
            ]
    \input{data-formatted/time-d4-verify}
    \input{data-formatted/time-d4-verify-onesided-only}
    \legend{
      \scriptsize \textsf{Full validation},
      \scriptsize \textsf{One-sided validation},
    }
    \addplot[mark=none, color=lightblue] coordinates{(0.1,0.01) (10000.0,1000.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,0.01) (10000.0,10000.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,0.1) (10000,100000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,1.0) (1000,100000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,10.0) (1000,1000000)};
    \node[left] at (axis cs: 0.9,0.12) {$0{.}1\times$};
    \node[left] at (axis cs: 0.09,0.12) {$1\times$};
    \node[left] at (axis cs: 0.09,1.2) {$10\times$};
    \node[left] at (axis cs: 0.09,12.0) {$100\times$};
    \node[left] at (axis cs: 0.09,120.0) {$1000\times$};

@@

\begin{figure}
\centering{%
\begin{tikzpicture}[scale = 0.90]
  \begin{axis}[mark options={scale=0.55},grid=both, grid style={black!10}, ymode=log,
      legend style={at={(0.22,0.98)}},
      legend cell align={left},
                              x post scale=1.8, y post scale=2.2,
                              xmode=log,xmin=0.01,xmax=1900,
                              xtick={0.01, 0.1,1.0,10,100,1000}, xticklabels={0.01, 0.1, 1.0, 10, 100,{1,000}},
                              ymin=0.01, ymax=1900,
                              ytick={0.01, 0.1,1.0,10,100,1000}, yticklabels={0.01, 0.1, 1.0, 10, 100, {1,000}},
                              xlabel={Structured generation + checking (seconds)}, ylabel={Monolithic generation + checking (seconds)},
            ]
    \input{data-formatted/time-structured+monolithic}
    \input{data-formatted/time-structured+monolithic_timeout}
    \addplot[mark=none, color=lightblue] coordinates{(1.0,0.01) (10000.0,100.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.1,0.01) (10000.0,1000.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,0.01) (10000.0,10000.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,0.1) (10000,100000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,1.0) (1000,100000)};
    \addplot[color=black,dashed] coordinates{(0.01,1500) (1500,1500)};
    \addplot[color=black,dashed] coordinates{(1500,0.01) (1500,1500)};
    \node[left] at (axis cs: 5.0,0.06) {$0{.}01\times$};
    \node[left] at (axis cs: 0.5,0.06) {$0{.}1\times$};
    \node[left] at (axis cs: 0.05,0.06) {$1\times$};
    \node[left] at (axis cs: 0.05,0.6) {$10\times$};
    \node[left] at (axis cs: 0.05,6.0) {$100\times$};

@@

Ran all benchmarks for which generation + checking could complete in 1000 seconds.  Total = 83.  All completed in less than 1000 seconds with structured.
Compare both runtimes and total number of clauses for structured vs. monolithic:
\begin{itemize}
\item Structured complete 83.
\item Monolithic completed  69, timed out on 14.
\item Of the 69 completed, 23 were faster with monolithic, 36 were slower (min ratio = 0.015, max ratio = 9.66, hmean = 0.40)
\item Of the 69 completed, 59 had fewer total clauses with monolithic, 10 had more (min ratio = 0.07, max ratio = 1.09, hmean = 0.49)
\end{itemize}

@@

%    \addplot[mark=none, color=black] coordinates{(0.01,1500) (1500,1500)};
%    \addplot[mark=none, color=black] coordinates{(1500,0.01) (1500,1500)};
    \addplot[mark=none, color=lightblue] coordinates{(10.0,0.01) (10000.0,10.0)};
    \addplot[mark=none, color=lightblue] coordinates{(1.0,0.01) (10000.0,100.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.1,0.01) (10000.0,1000.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,0.01) (10000.0,10000.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,0.1) (10000,100000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,1.0) (1000,100000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,10.0) (1000,1000000)};
    \node[left] at (axis cs: 50,0.06) {$0{.}001\times$};
    \node[left] at (axis cs: 5.0,0.06) {$0{.}01\times$};
    \node[left] at (axis cs: 0.5,0.06) {$0{.}1\times$};
    \node[left] at (axis cs: 0.05,0.06) {$1\times$};
    \node[left] at (axis cs: 0.05,0.6) {$10\times$};
    \node[left] at (axis cs: 0.05,6.0) {$100\times$};
    \node[left] at (axis cs: 0.05,60.0) {$1000\times$};

@@

We encountered a significant number of errors with both
\textsc{nnf2trace} and \textsc{sharptrace}.  Based on the error
messages, these appear to be mostly due to features of some of the
d-DNNF graphs they had not anticipated.  We provided a bug report to
the authors, but they were unable to fix them on short notice.

@@

Overall, we can summarize the results as:
\begin{itemize}
\item 84 total data points
\item 63 were completed by both programs in under 1000 seconds
\item Of those 20 were faster with MICE, 43 with CRAT
\item 6 completed with CRAT under 1000 seconds but exceeded that threshold for MICE
\item 1 completed with MICE under 1000 seconds but exceeded that threshold for CRAT
\item 14 completed with CRAT but generated errors for MICE.
\end{itemize}

>>>>

@@

\begin{figure}
\centering{%
\begin{tikzpicture}[scale = 0.90]
  \begin{axis}[mark options={scale=0.55},grid=both, grid style={black!10}, ymode=log,
      legend style={at={(0.30,0.96)}},
      legend cell align={left},
                              x post scale=1.8, y post scale=2.4,
                              xmode=log,xmin=0.01,xmax=4000,
                              xtick={0.01, 0.1,1.0,10,100,1000,10000}, xticklabels={0.01, 0.1, 1.0, 10, 100, {1,000}, {10,000}},
                              ymin=0.01, ymax=29000,
                              ytick={0.01, 0.1,1.0,10,100,1000,10000,100000}, yticklabels={0.01, 0.1, 1.0, 10, 100, {1,000},{10,000},{100,000}},
                              xlabel={D4 runtime (seconds)}, ylabel={CRAT generation and checking runtime (seconds)},
%                              title={D4 Defining Clause Generation}
            ]
    \input{data-formatted/time-d4-verify}
    \input{data-formatted/time-d4-verify-onesided-only}
    \input{data-formatted/time-d4-failure}
    \legend{
      \scriptsize \textsf{Full validation},
      \scriptsize \textsf{One-sided validation},
      \scriptsize \textsf{Resource limit exceeded},
    }
    \addplot[mark=none, dashed] coordinates{(0.01,25000) (4000, 25000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.1,0.01) (10000.0,1000.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,0.01) (10000.0,10000.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,0.1) (2000,20000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,1.0) (200, 20000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,10.0) (20, 20000)};
    \node[left] at (axis cs: 0.9,0.12) {$0{.}1\times$};
    \node[left] at (axis cs: 0.09,0.12) {$1\times$};
    \node[left] at (axis cs: 0.09,1.2) {$10\times$};
    \node[left] at (axis cs: 0.09,12.0) {$100\times$};
    \node[left] at (axis cs: 0.09,120.0) {$1000\times$};

@@

@@

\begin{figure}
\centering{%
\begin{tikzpicture}[scale = 0.90]
  \begin{axis}[mark options={scale=0.55},grid=both, grid style={black!10}, ymode=log,
      legend style={at={(0.22,0.98)}},
      legend cell align={left},
                              x post scale=1.8, y post scale=2.2,
                              xmode=log,xmin=0.01,xmax=1900,
                              xtick={0.01, 0.1,1.0,10,100,1000}, xticklabels={0.01, 0.1, 1.0, 10, 100,{1,000}},
                              ymin=0.01, ymax=1900,
                              ytick={0.01, 0.1,1.0,10,100,1000}, yticklabels={0.01, 0.1, 1.0, 10, 100, {1,000}},
                              xlabel={Structured generation + checking (seconds)}, ylabel={Monolithic generation + checking (seconds)},
            ]
    \input{data-formatted/time-structured+monolithic}
    \input{data-formatted/time-structured+monolithic_timeout}
    \addplot[mark=none, color=lightblue] coordinates{(1.0,0.01) (1000.0,10.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.1,0.01) (1000.0,100.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,0.01) (1000.0,1000.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,0.1) (100, 1000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,1.0) (10, 1000)};
    \addplot[color=black,dashed] coordinates{(0.01,1500) (1500,1500)};
    \addplot[color=black,dashed] coordinates{(1500,0.01) (1500,1500)};
    \node[left] at (axis cs: 5.0,0.06) {$0{.}01\times$};
    \node[left] at (axis cs: 0.5,0.06) {$0{.}1\times$};
    \node[left] at (axis cs: 0.05,0.06) {$1\times$};
    \node[left] at (axis cs: 0.05,0.6) {$10\times$};
    \node[left] at (axis cs: 0.05,6.0) {$100\times$};

@@

Ran all benchmarks for which generation + checking could complete in 1000 seconds.  Total = 82.  All completed in less than 1000 seconds with structured.
Compare both runtimes and total number of clauses for structured vs. monolithic:
\begin{itemize}
\item Structured complete 82.
\item Monolithic completed  69, timed out on 13.
\item Of the 69 completed, 23 were faster with monolithic, 36 were slower (min ratio = 0.015, max ratio = 9.66, hmean = 0.40)
\item Of the 69 completed, 59 had fewer total clauses with monolithic, 10 had more (min ratio = 0.07, max ratio = 1.09, hmean = 0.49)
\end{itemize}

@@

%    \addplot[mark=none, color=black] coordinates{(0.01,1500) (1500,1500)};
%    \addplot[mark=none, color=black] coordinates{(1500,0.01) (1500,1500)};
    \addplot[mark=none, color=lightblue] coordinates{(10.0,0.01) (1000.0,1.0)};
    \addplot[mark=none, color=lightblue] coordinates{(1.0,0.01) (1000.0,10.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.1,0.01) (1000.0,100.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,0.01) (1000.0,1000.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,0.1) (100,1000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,1.0) (10,1000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,10.0) (1,1000)};
    \node[left] at (axis cs: 50,0.06) {$0{.}001\times$};
    \node[left] at (axis cs: 5.0,0.06) {$0{.}01\times$};
    \node[left] at (axis cs: 0.5,0.06) {$0{.}1\times$};
    \node[left] at (axis cs: 0.05,0.06) {$1\times$};
    \node[left] at (axis cs: 0.05,0.6) {$10\times$};
    \node[left] at (axis cs: 0.05,6.0) {$100\times$};
    \node[left] at (axis cs: 0.05,60.0) {$1000\times$};

@@

We encountered a significant number of errors with both
\textsc{nnf2trace} and \textsc{sharptrace}.  Based on the error
messages, these appear to be mostly due to features of some of the
CNF files or d-DNNF graphs they had not anticipated.  We provided a bug report to
the authors, but they were unable to fix them on short notice.

@@

Overall, we can summarize the results as:
\begin{itemize}
\item 84 total data points
\item 67 were completed by both programs in under 1000 seconds
\item Of those 21 were faster with MICE, 46 with CRAT
\item 7 completed with CRAT under 1000 seconds but exceeded that threshold for MICE
\item 1 completed with MICE under 1000 seconds but exceeded that threshold for CRAT
\item 9 completed with CRAT but generated errors for MICE.
\end{itemize}
